### 2018-06-14 chapters/introduction/junk.tex
more texts.

Hold-Geoffroy \etal~\cite{hold2017perceptual}
improve~\cite{workman2016horizon}'s results by using deeper networks,
and introduce a new perceptual measure for horizon line detection.
Unlike the previous work that mainly focus on horizon line detection,
Zhang \etal~\cite{zhang2018dominant} propose to detect the dominant
vanishing point using deep neural networks.

\brief{
Geometric information can help training neural network to extract
high-level information.
}

\todo{re-edit}
The use of overhead imagery has been proved useful for image
understanding.
Several methods have been recently proposed to jointly reason about
co-located aerial and ground image pairs. Luo
\etal~\cite{luo2008event} demonstrate that aerial imagery can aid
in recognizing the visual content of a geo-tagged ground image.
M{\'a}ttyus \etal~\cite{mattyus2016hd} perform joint inference over
monocular aerial imagery and stereo ground images for fine-grained
road segmentation. Wegner \etal~\cite{wegner2016cataloging} build a
map of street trees. Given the horizon line and the camera intrinsics,
Ghouaiel and Lef{\`e}vre~\cite{ghouaiel2016coupling} transform
geo-tagged ground-level panoramas to a top-down view to enable
comparisons with aerial imagery for the task of change detection.
Recent work on cross-view image
geolocalization~\cite{lin2013cross,lin2015learning,workman2015geocnn,workman2015wide}
 has shown that convolutional neural
networks are capable of extracting features from aerial imagery
that can be matched to features extracted from ground imagery.
Vo \etal~\cite{vo2016localizing} extend this line of work,
demonstrating improved geolocalization performance by applying an
auxiliary loss function to regress the ground-level camera orientation
with respect to the aerial image. To our knowledge, our work is the
first work to explore predicting the semantic layout of a ground
image from an aerial image.