### 2018-06-27 chapters/introduction/junk.tex
more

Hold-Geoffroy \etal~\cite{hold2017perceptual}
improve~\cite{workman2016horizon}'s results by using deeper networks,
and introduce a new perceptual measure for horizon line detection.
Unlike the previous work that mainly focus on horizon line detection,
Zhang \etal~\cite{zhang2018dominant} propose to detect the dominant
vanishing point using deep neural networks.

\brief{
Geometric information can help training neural network to extract
high-level information.
}

\todo{re-edit}
The use of overhead imagery has been proved useful for image
understanding.
Several methods have been recently proposed to jointly reason about
co-located aerial and ground image pairs. Luo
\etal~\cite{luo2008event} demonstrate that aerial imagery can aid
in recognizing the visual content of a geo-tagged ground image.
M{\'a}ttyus \etal~\cite{mattyus2016hd} perform joint inference over
monocular aerial imagery and stereo ground images for fine-grained
road segmentation. Wegner \etal~\cite{wegner2016cataloging} build a
map of street trees. Given the horizon line and the camera intrinsics,
Ghouaiel and Lef{\`e}vre~\cite{ghouaiel2016coupling} transform
geo-tagged ground-level panoramas to a top-down view to enable
comparisons with aerial imagery for the task of change detection.
Recent work on cross-view image
geolocalization~\cite{lin2013cross,lin2015learning,workman2015geocnn,workman2015wide}
 has shown that convolutional neural
networks are capable of extracting features from aerial imagery
that can be matched to features extracted from ground imagery.
Vo \etal~\cite{vo2016localizing} extend this line of work,
demonstrating improved geolocalization performance by applying an
auxiliary loss function to regress the ground-level camera orientation
with respect to the aerial image. To our knowledge, our work is the
first work to explore predicting the semantic layout of a ground
image from an aerial image.

The study of understanding images has been around for half a century.
A well known anecdote in computer vision community happened in 1960s,
when the MIT professor Dr. Marvin Minsky assigned a summer programming
project to his undergraduates, to recognize objects in a
scene~\cite{boden2006mind}. 
%
Similar to the story of the famous ``Four Color Conjecture'', where
one of the world's renowned mathematicians Dr. Hermann Minkowski
claimed in his class that he was able to solve it before the class
ends.  
However, Not until one hundred years later since it was first
proposed, the ``Four Color Conjecture'' has been stamped as theorem by
the continuous efforts of several generations of talented
mathematicians.  While on the other side, Dr. Marvin Minsky's summer
program still remains open.

To achieve this
goal, researchers adopt a divide-and-conquer strategy, that is
dividing image understanding into various sub-problems like object
recognition, semantic segmentation, and pose estimation \etc.  During
past decades, a lot of methods are developed to solve these
sub-problems. Many of them adopted a bottom-up (\todo{citations})
workflow, where basic level cues play important roles
(\todo{examples}). Among these basic cues, camera geometry is
particularly useful for some computer vision tasks like facade
detection, indoor layout estimation~\cite{ren2016coarse}.

In the camera pin-hole model, camera geometric parameters consists of
the camera intrinsics and the extrinsics. The intrinsic parameters of
a camera encode information of the camera itself such as focal length,
principal point \etc; the extrinsic parameters are referred to as the
external states of the camera, such as the camera poses and
offset/location under some coordinate system. In this work, our study
of the camera geometry mainly includes the extrinsic like camera pose
and the geo-orientation/geo-localization of the camera.

\sepline

Among algorithms that are developed to address these sub-tasks, deep
neural networks (DNNs) rise as stars in recent years. Methods that
uses DNNs achieve state-of-the-art performances in many computer
vision areas (\todo{citations}). As parametric algorithms, deep neural
networks require training for their parameters. Methods to train DNNs
include supervised algorithms, unsupervised algorithms, and
reinforcement learning. Supervised algorithms are the most popular and
commonly applied in many computer vision areas, thanks to their good
performances and the fact that they are easy to implement. However,
supervised learning approaches are sensitive to the quality of data
annotations. In some applications like overhead image segmentation and
transient attributes estimation, data with noise-free annotation is
either limited in quantity or technically difficult to acquire.
In this thesis, Our methods focus on lavaging the camera geographic
information to deal with the challenges caused by the lack of
noise-free annotations during DNN training. The insight is
that the camera geographic information encodes the semantic
information because it is related to the appearances of
the scenes.

The process of estimating the camera geographic information is
referred to as geo-calibration. The goal is to estimate the geographic
location and the geo-orientation of the given camera. 
Compared to the common camera calibration, geo-calibration focuses on
calibrating the camera in the geographic coordinates.

Geographic
information reveals the scene appearances

\sepline

Many research has been done to detect the camera
geometry from input images. \todo{citation and examples}. Most of the
previous work optimize for a limited set of camera calibration
parameters only. However, our research found that there exists an
effective way of optimizing multiple camera calibration parameters at
simultaneously (refer to \chapref{mcmc}). Our algorithm uses the Monte
Carlo Markov Chain (MCMC) method to sample the possible solutions that
minimize the objective function, which measures the ``matchness''
between the observation (input image) and the camera calibration.

Among the most influential approaches for camera geometry estimation,
most methods are based on traditional statistics diagram. Not until a
fewer years ago, machine learning based methods gradually caught
people's attention in this area, thanks for the popularity of deep
neural network. Neural networks were first proposed in 80th
(\todo{citation}), The back-forward method developed by Hinton
\todo{citation} makes the effective training possible. However, neural
networks are generally thought by researchers merely toy that can only
solve a very limited set of tasks. It is the develop of computational
resources and the big collection of annotated data that enables the
neural networks to grow deeper and eventually brings it to the stage of
industrial applications.

In computer vision, low-level elements like edges, line segments,
super-pixels and vanishing points play important roles in many
applications like image measuration, facade detection, and
geo-localization.
Though the CNNs are good at extracting high-level information from
images, more and more researches start focusing on using CNN to
extract low-level these elements.
%
Lee \etal~\cite{lee2017semantic} propose to detect semantic line
segments, which are able to segment the scene semantically, using
neural network. Maninis \etal~\cite{maninis2016convolutional} and Yang
\etal~\cite{yang2016object} explore to detect object contours with
CNNs.
%
Aside from edges and contours, another category of low-level
information that the CNNs can detect is the camera geometry, like
camera location, vanishing points, and horizon lines.
%
In the area of vanishing point detection, we are among the first
researches that use deep neural networks to solve the problem. The
deep neural networks prove to be a good tool to extract high-level
information that offer great help to conquer the problems in camera
geometry estimation. \todo{many citations}.

In our previous work, we have shown that deep neural networks are
powerful tools for camera geometry estimation as well. In fact, the
camera geometry can also assist learning high-level image semantics.
Hoiem \etal~\cite{hoiem2008putting} demonstrate that taking camera
perspective (horizon line) into consideration improves the performance
for object detection. Godard \etal~\cite{godard2017unsupervised} learn
to predict the scene depth from a single image using left-right
consistency, which is also an application of camera geometry.
\todo{more citations about the projective loss}.

Another example of camera geometry boosting the learning project is to
exploit the overhead imagery. Using the geometric locations of the
ground-view images we can download the corresponding overhead images
at the same locations through public online resources like bingmap
\todo{cite}. We refer to these image pairs as ``cross-view image
pairs''.

\todo{copy from scott} Efforts have been made to
automate overhead image analysis. As early as
1970~\cite{idelsohn1970learning} methods were introduced for
classifying terrain types from a single overhead image, with the goal
of automatically generating terrain maps.  Similarly, in 1976 Bajcsy
et al.~\cite{bajcsy1976computer} described a system for recognizing
roads, intersections, and other road-like objects in overhead imagery.
However, as overhead imaging differs drastically from ground-level
imaging, the majority of techniques that have been developed have
occurred independently and in task-specific ways~\cite{Rozen}.
%\brief{overhead image grows fast}.

Our work demonstrates that we are able to transfer the learned semantic
knowledge from the ground image domain to the overhead image domain,
given the camera geometric information of these two kind of images. In
our algorithm, the camera geometry helps to collect the ``cross-view
image pairs'' and aligned the orientation directions of images in a
pair.

Camera geometry can also be used as weak annotations during training
of the networks for image understanding.  One of the key to the
success of machine learning approaches are the large quantity of
annotated data. However, fully annotated imagery consists of only a
tiny part of the image resources available online. The idea is to
exploit other meta-info that are automatically recorded/captured by
devices to train the ML models. One of the meta-info potentially
helpful for learning is camera geometry.  We demonstrate that the
camera geometry can be used as weak ground truth for supervision
problem. By learning the capture time and the geolocation of the
images simultaneously, our network is able capture the geo-temporal
related high-level features.