### 2018-08-03 dissertation.tex
introduction first para

\documentclass[final]{ukthesis}

\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{paralist}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{import}
\usepackage{xspace}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{memhfixc}
\usepackage{algorithm2e, algorithmic}
\usepackage{soul}
\usepackage{multirow}
\usepackage{miscstyle}

\usepackage[pdfauthor={Menghua Zhai},
            pdftitle={dissertation},
            bookmarks=false,
            bookmarksnumbered=true,
            bookmarksopen=true,
            breaklinks=true,
            letterpaper=true,
            linktocpage,
            pagebackref=true]{hyperref}

\begin{document}
\emergencystretch 3em

\author{Menghua Zhai}
%\title{Building a Probabilistic Model with Deep Neural Networks for Camera Geo-Calibration}
%\title{Statistical Models for Camera Geo-Calibration}
\title{Deep Probabilistic Models for Camera Geo-Calibration}

\abstract
{ \SingleSpacing
The ultimate goal of image understanding is to transfer the visual
images into numerical or symbolic descriptions of the world
that are helpful for decision making.
%
%In general, understanding images is not trivial for machines.
%Under different scenarios, researchers divide image
%understanding into different computer vision tasks. For example,
%pedestrian detection is essential to autonomous driving, and image
%retrieval to an image searching engine.
%
Knowing when, where, and in which direction picture was taken, the
task of geo-calibration makes it possible to use imagery to understand
the world and how it changes in time. Current models for
geo-calibration are mostly deterministic, namely they output fixed
solutions. However, deterministic models can not model the
uncertainties that are inherent when the scene of the image is ambiguous.
Furthermore, their outputs are not very friendly to algorithms that
take probabilistic inputs. To address these limitations, we
propose a probabilistic model for camera geo-calibration using deep
neural networks.
%
While our primary contribution is geo-calibration, We also show that
learning to geo-calibrate a camera is helpful for high-level feature
extraction.
%
%Our discoveries can provide researchers new insights for both camera
%geo-calibration and DNN training.
}
\keywords{computer vision, geo-calibration, deep neural networks}

\advisor{Nathan Jacobs}
\dgs{Miroslaw Truszczynski}

\frontmatter
\maketitle

%\begin{acknowledgments}
	%Acknowledge people/things here
%\end{acknowledgments}

%\begin{dedication}
%	Dedicated to things (optional)
%\end{dedication}

\tableofcontents\clearpage
%\listoffigures\clearpage
%\listoftables\clearpage

\mainmatter

\inject{introduction}

%\inject{mcmc}
%\inject{fasthorizon}
%\inject{crosstransf}
%\inject{whenwhere}

%\inject{discussion}

\backmatter

\bibliographystyle{plain}
\bibliography{biblio}

%\inject{vita}

\end{document}