### 2018-09-17 chapters/discussion/discussion.tex
minor change

\chapter{Discussion}
\label{chap:discussion}

The goal of geo-calibration is to learn the camera pose, location and
the time when the image was captured.
Our thesis focused on developing deep geo-calibration algorithms for
image understanding.
Compared to previous work, our approaches output probabilistic
predictions that are able to handle the uncertainty of the scene
better. Furthermore, we also show that
learning to geo-calibrate a camera allows us to implicitly learn to
understand the content of the scene.

In \chapref{mcmc}, we proposed a flexible model to compute the
probability over all geo-calibration parameters. In the high
dimensional space of geo-calibration parameters, we defined a score
function for every point, to measure how well the
geo-calibration parameters match the content of the input image.
The score function is a weighted summation of a series of
constraint functions. It is proportional to the probability density
function of the distribution. Since the analytic solution to the
integral of the score function does not exist, we approximated the
probability density function using MCMC method. Our model is flexible
to any constraint function for geo-calibration parameters. We also
demonstrated in our experiment that more accurate predictions can be
obtained if the score function consists of more constraints.

In \chapref{fasthorizon}, we focused on developing constraint function
for the camera pose. Assuming the focal length is known, the camera
roll and pitch angle can be derived from the horizon line position.
Our algorithm identifies the horizon line from the input image
satisfying the Manhattan world assumption. Compared to the previous
work, which applies a bottom-up approach where the horizon line is
estimated after the vanishing point identification, we proposed a
novel horizon-first approach where finding horizon line candidates is
prior to identifying vanishing points.  As the first step, we trained
a CNN to estimate the probability distribution over the horizon line.
Then we sample horizon line candidates from the distribution. For each
horizon candidate, we detect likely vanishing points on it. Compared
to previous work which searches for vanishing points on the 2D image
frame, our method is able to accelerate the searching process by
reducing the searching space to 1D. The probability of the horizon
line is measured by how convincing are the vanishing points found on
it.

In \chapref{whenwhere}, we constructed the constraint function for 
the image capture time and the camera location. In order to achieve
this goal, we trained a CNN on a large dataset that consists of
millions webcam frames and smart phone photos. The network takes image
as input and output the discrete probability over location and time.
In this work, we took the advantage that geo-tagged and
temporal-tagged images are common nowadays to build the training set.
We demonstrate that geo-calibrating the camera helps learn image
representations useful for image understanding.

In \chapref{crosstransf}, we exploit overhead imagery to construct
the constraint function for the camera orientation and location.
We trained a CNN that can estimate the semantic layout of the ground
view image from the overhead image. During training, we learn the
pixel-to-pixel correspondences from the aerial image to the ground. We
project the semantic layout the of the overhead image to the ground
view with the learned correspondence.
When computing the output of the constraint function, we feed the
aerial image of the gien location and orientation to the network. By
comparing the aerial-to-ground with the real layout of the ground
image, we measure the consistency between the two and use
it as the value of our constraint function.
We also demonstrate the efficiency of our method for CNN pre-training
for aerial image segmentation.

This thesis proposed a probabilistic model for image
geo-calibration.
The key to the success of this model is to find constraint
functions that describe the relationship between geo-calibration
parameters and the image content. We developed constraint
functions for different set of parameters that can fit to the model.
Furthermore, we also show that learning to geo-calibrate cameras helps learning
useful features for image understanding.
There are several possible future research directions for extending
this work, including: exploiting map data in cross-view learning,
replacing the CNN networks in our algorithms with Bayesian neural
networks (BNNs).
Overall, we hope our work can provide new ways for geo-calibration and
pre-task training.