### 2018-09-18 chapters/discussion/discussion.tex
minor change

\chapter{Discussion}
\label{chap:discussion}

The goal of geo-calibration is to learn the camera pose, location and
the time when the image was captured.
Our thesis focused on developing deep geo-calibration algorithms for
image understanding.
Compared to previous work, our approaches output probabilistic
predictions that are able to handle the uncertainty of the scene
better. Furthermore, we also show that
learning to geo-calibrate a camera allows us to implicitly learn to
understand the content of the scene.

In \chapref{mcmc}, we proposed a flexible model to compute the
probability over all geo-calibration parameters. In the high
dimensional space of geo-calibration parameters, we defined a score
function for every point, to measure how well the
geo-calibration parameters match the content of the input image.
The score function is a weighted summation of a series of
constraint functions. It is proportional to the probability density
function of the distribution. Since the analytic solution to the
integral of the score function does not exist, we approximated the
probability density function using MCMC method. Our model is flexible
to any constraint function for geo-calibration parameters. We also
demonstrated in our experiment that more accurate predictions can be
obtained if the score function consists of more constraints.

In \chapref{fasthorizon}, we focused on developing constraint function
for the camera pose. Assuming the focal length is known, the camera
roll and pitch angle can be derived from the horizon line position.
Our algorithm identifies the horizon line from the input image
satisfying the Manhattan world assumption. Compared to the previous
work, which applies a bottom-up approach where the horizon line is
estimated after the vanishing point identification, we proposed a
novel horizon-first approach where finding horizon line candidates is
prior to identifying vanishing points.  As the first step, we trained
a CNN to estimate the probability distribution over the horizon line.
Then we sample horizon line candidates from the distribution. For each
horizon candidate, we detect likely vanishing points on it. Compared
to previous work which searches for vanishing points on the 2D image
frame, our method is able to accelerate the searching process by
reducing the searching space to 1D. The probability of the horizon
line is measured by how convincing are the vanishing points found on
it.

In \chapref{whenwhere}, we developed the constraint function for 
the image capture time and the camera location. 
We trained a CNN on a large dataset to learn to predict the discrete
probabilities over location and time given the input image.
Unlike full supervision methods which require tedious work for manual
annotations, our training dataset consists of millions webcam
frames and smart phone photos with free time/location tags.
In the experiments, we demonstrated that geo-calibrating the camera
also helps learn useful image representations for image understanding.

In \chapref{crosstransf}, we exploited overhead imagery to construct
the constraint function for the camera orientation and location.
We trained a CNN that can estimate the semantic layout of the ground
view image from the overhead image. During training, the CNN learns the
pixel-to-pixel correspondences from the aerial image to the ground
image. 
It then projects the semantic layout the of the overhead image to the
ground view according to the learned correspondence.
When computing the output of the constraint function, we feed the
aerial image of given location and orientation to the network. By
comparing the aerial-to-ground layout with the layout computed from
the ground image, we measure the consistency between these two and use
consistency value as the output of the constraint function. 
In our experiments, we also demonstrated the efficiency of our method
for pre-training CNNs for aerial image segmentation.

This thesis proposed a probabilistic model for image geo-calibration.
The key to the success of this model is to find constraint functions
that describe the relationship between geo-calibration parameters and
the image content. We developed constraint functions for different set
of parameters that can fit to the model.  Furthermore, we also show
that learning to geo-calibrate cameras helps learning useful features
for image understanding.  There are several possible future research
directions for extending this work, including: exploiting map data in
cross-view learning, replacing the CNN networks in our algorithms with
Bayesian neural networks (BNNs).  Overall, we hope our work can
provide new ways for geo-calibration and pre-task training.